Global adjoint tomography -- Model GLAD-M25\footnote{This paper is dedicated to Dimitri Komatitsch, computational geophysicist extraordinaire, who passed away unexpectedly shortly before its submission.}

### 2019-01-29 main.tex
Updates from ShareLaTeX

@@

To illustrate which seismic phases end up being picked,
Fig.~\ref{fig:window_density} shows a time-distance plot
for vertical component 17--40~s body waves in which the Pyflex window count is used to identify parts of seismograms being assimilated.
A total of 4,363,719 time windows were picked in this category.
The figure nicely highlights all the well known vertical-component body traveltime branches.
The maximum window count is 9,931 in the 134--135$^\circ$ distance and 1375--1,376~s time bin.

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{figures/window_colorbar_linear_Z.pdf}
  \caption{Time-distance plot for vertical component 17--40~s body waves showing a measurement window density map.
  A total of 4,363,719 time windows were discretized into 1~degree epicentral distance and 1~s time bins. Major traveltime branches are labeled. Note that the plot contains data from events with a wide variety of depths, so the branches are for identification purposes only.
  }
  \label{fig:window_density}
\end{figure}

@@

There are more than ten stages in the adjoint tomography workflow,
and each stage involves thousands of small tasks.
For every model iteration,
we performed up to 1,480 forward and adjoint simulations, each involving hundreds of
compute cores and graphics cards, as well as heavy I/O.
The workflow is prone to human error and hardware failure, making it fragile~~\citep{Lefebvre2018}.
For these reasons, we hardened the workflow by taking advantage of modern
workflow management software.
We selected RADICAL-SAGA and RADICAL-EnTK as our
workflow management engines, and developed complementary seismic tomography
workflow tools~\citep{EnTK2017}.
The workflow engine can automatically detected job failures both from the
HPC system and via user-defined functions. This enables us to keep track of
all tasks and semi-automatically resubmit jobs if necessary.
Given that most HPC system time is spent waiting in the job queue, automatic
failure detection and relaunching greatly shortens the overall time to solution.

@@

Fig.~\ref{fig:P_phase_hist} shows histograms of 17--40~s P-wave traveltime anomalies on the vertical and radial components.
Both components exhibit consistent reductions in terms of both mean and variance. 
These arrivals --- almost 900,000 measurements in the 17--40~s and 40--100~s period bands, roughly~$\sim8\%$ of the database --- are of great importance, because they help constrain the compressional wavespeed structure in the model.

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/dt_histogram_P.pdf}
  \caption{Histograms of traveltime measurements of 17--40~s P arrivals on the vertical and radial for S362ANI combined with CRUST2.0 (M00, Black), GLAD-M15 (M15, Green), and GLAD-M25 (M25, Red). Other settings are the same as in Fig.~\ref{fig:phase_hist}. {\color{red}Wenjie: remove 40--100~s.}}
  \label{fig:P_phase_hist}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/dt_histogram_360.pdf}
  \caption{Histograms of phase measurements for the 360 events in the held-out database.}
  \label{fig:phase_hist_360}
\end{figure}

@@

To assess the accuracy of the expansion,
we calculate the model norm in the GLL basis,
\begin{equation}
    ||m_\text{\,GLL}||^2=\int_V m_\text{\,GLL}^2(\mathbf{x})\,\mathrm{d}^3\mathbf{x}=
    \sum_{\mathrm{elem}}\sum_{\alpha,\beta,\gamma}\omega_\alpha\,\omega_\beta\,\omega_\gamma\,(m^{\alpha\beta\gamma})^2\,J^{\alpha\beta\gamma}
    \quad ,
\end{equation}
and compare it to the model norm in the spherical harmonic basis,
\begin{equation}
    ||m_\text{\,SH}||^2=\int_V m_\text{\,SH}^2(\mathbf{x})\,\mathrm{d}^3\mathbf{x}=
    \sum_{\ell = 0}^L(2\ell+1)\,\sigma_\ell^2
    \quad ,
\end{equation}
by calculating the error
\begin{equation}
    \epsilon = \log\,\frac{||m_\text{\,SH}||^2}{||m_\text{\,GLL}||^2}
    \quad .
\end{equation}

>>>>

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{figures/window_colorbar_linear_Z.pdf}
  \caption{Time-distance plot for vertical component 17--40~s body waves showing a measurement window density map.
  A total of 4,363,719 time windows were discretized into 1~degree epicentral distance and 1~s time bins.
  The maximum window count is 9,931 in the 134--135$^\circ$ distance and 1375--1,376~s time bin.
  Major traveltime branches are labeled. Note that the plot contains data from events with a wide variety of depths, so the branches are for identification purposes only.
  }
  \label{fig:window_density_Z}
\end{figure}

@@

To illustrate which seismic phases end up being picked,
Figs.~\ref{fig:window_density_Z}, \ref{fig:window_density_R}, and \ref{fig:window_density_T} show time-distance plots
for vertical, radial, and transverse component 17--40~s body waves,
respectively,
in which the Pyflex window count is used to identify parts of seismograms being assimilated.
The figures nicely highlight all the well known body traveltime branches.
The Pyflex window selection algorithm is currently configured not to pick body waves after the arrival of the surface waves, which is why the PcP and ScS branches are missing at shorter epicentral distances.
This is something we will reconsider in future applications, perhaps by introducing aspects of machine learning.

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{figures/window_colorbar_linear_R.pdf}
  \caption{Same as Fig.~\ref{fig:window_density_Z}, except for radial component 17--40~s body waves.
   A total of 2,868,252 time windows were discretized into 1~degree epicentral distance and 1~s time bins.
   The maximum window count is 7,542.
   }
  \label{fig:window_density_R}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{figures/window_colorbar_linear_T.pdf}
  \caption{Same as Fig.~\ref{fig:window_density_Z}, except for transverse component 17--40~s body waves.
   A total of 2,151,856 time windows were discretized into 1~degree epicentral distance and 1~s time bins.
   The maximum window count is 7,208.
   Only SH travetime branches are labeled.}
  \label{fig:window_density_T}
\end{figure}

@@

There are more than ten stages in the adjoint tomography workflow,
and each stage involves thousands of small tasks.
For every model iteration,
we performed up to 1,480 forward and adjoint simulations, each involving hundreds of
compute cores and graphics cards, as well as heavy I/O.
The workflow is prone to human error and hardware failure, making it fragile~~\citep{Lefebvre2018}.
For these reasons, we hardened the workflow by taking advantage of modern
workflow management software.
We selected EnTK as our
workflow management engine~\citep{EnTK2017}, and developed complementary seismic tomography
workflow tools.
The workflow engine can automatically detected job failures both from the
HPC system and via user-defined functions. This enables us to keep track of
all tasks and semi-automatically resubmit jobs if necessary.
Given that most HPC system time is spent waiting in the job queue, automatic
failure detection and relaunching greatly shortens the overall time to solution.

@@

Fig.~\ref{fig:P_phase_hist} shows histograms of 17--40~s P-wave traveltime anomalies on the vertical and radial components.
Both components exhibit consistent reductions in terms of both mean and variance. 
Collectively, the numerous compressional wave arrivals identified in Figs.~\ref{fig:window_density_Z} and~\ref{fig:window_density_R} are of great importance, because they help constrain the compressional wavespeed structure in the model.

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/dt_histogram_P.pdf}
  \caption{Histograms of 17--40~s P-wave traveltime anomalies of the vertical and radial components for models S362ANI combined with CRUST2.0 (M00, Black), GLAD-M15 (M15, Green), and GLAD-M25 (M25, Red).
  }
  \label{fig:P_phase_hist}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/dt_histogram_360.pdf}
  \caption{Same as Fig.~\ref{fig:phase_hist}, except for the held-out database.
  }
  \label{fig:phase_hist_360}
\end{figure}

P phase measurements were also extracted and compared in Fig.~\ref{fig:phase_hist_360_P} for 3 models. The behavior is consistent as we expected.

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/dt_histogram_360_P.pdf}
  \caption{Same as Fig.~\ref{fig:P_phase_hist}, except for the held-out database.}
  \label{fig:phase_hist_360_P}
\end{figure}

@@

To assess the accuracy of the expansion,
we may calculate the model norm in the GLL basis,
\begin{equation}
    ||m_\text{\,GLL}||^2=\int_V m_\text{\,GLL}^2(\mathbf{x})\,\mathrm{d}^3\mathbf{x}=
    \sum_{\mathrm{elem}}\sum_{\alpha,\beta,\gamma}\omega_\alpha\,\omega_\beta\,\omega_\gamma\,(m^{\alpha\beta\gamma})^2\,J^{\alpha\beta\gamma}
    \quad ,
\end{equation}
and compare it to the model norm in the spherical harmonic basis,
\begin{equation}
    ||m_\text{\,SH}||^2=\int_V m_\text{\,SH}^2(\mathbf{x})\,\mathrm{d}^3\mathbf{x}=
    \sum_{\ell = 0}^L(2\ell+1)\,\sigma_\ell^2
    \quad ,
\end{equation}
by calculating the error
\begin{equation}
    \epsilon = \log\,\frac{||m_\text{\,SH}||^2}{||m_\text{\,GLL}||^2}
    \quad .
\end{equation}
